---
title: "Bivariate Viz"
author: Phoebe Pan
date: 4/24/2025
format: 
  html:
    code-fold: true
    theme: cosmo
    toc: true
    number-sections: true
    warning: false
---

This section showcases my best bivariate visualization, developed and improved throughout the semester as I learned principles of effective data visualization and design. The goal was to clearly communicate the relationship between two variables using clean, informative, and professional plots.

The data comes from two sources: the 2020 U.S. county-level election results (election_2020_county.csv) and weather data from three locations (weather_3_locations.csv), both provided through the course website at mac-stat.github.io. These datasets allow for exploration of how weather patterns may relate to regional political preferences or other contextual variables.

```{r setup, include=FALSE}
library(ggplot2)
library(readr)
library(tidymodels)
library(tidyverse)
```

```{r, include=FALSE}
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv")
```

```{r}
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill") +
  scale_fill_manual(
    name = "2020 Winner",
    values = c("dem" = "#5DADE2", "repub" = "#F1948A"),
    labels = c("Democrat", "Republican")
  ) +
  labs(
    title = "2020 Election Outcomes by Historical State Voting Patterns",
    x = "Historical Voting Trend",
    y = "Proportion of Counties",
    caption = "Source: ICA"
  ) +
  theme_minimal(base_size = 14)

```
```{r}
ggplot(weather, aes(x = temp3pm, fill = location)) +
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Distribution of 3PM Temperatures by Location",
    x = "Temperature at 3PM (°C)",
    y = "Density",
    fill = "Location",
    caption = "Source: ICA"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")
```


