---
title: "Quadvariate Viz"
author: Phoebe Pan
date: 4/24/2025
format: 
  html:
    code-fold: true
    theme: cosmo
    toc: true
    number-sections: true
    warning: false
---

This section showcases my best quadvariate visualization, developed and refined throughout the semester as I learned how to encode and communicate multiple variables within a single cohesive plot. The goal was to represent four distinct dimensions of data simultaneously using effective combinations of position, color, shape, and grouping.

The data comes from the weather_3_locations.csv dataset, available at mac-stat.github.io, which contains daily weather observations—including temperature, humidity, wind speed, and precipitation—for three different locations. This dataset provided an excellent opportunity to explore complex interactions across time, space, and meteorological variables in a meaningful visual format.

```{r, include=FALSE}
library(ggplot2)
library(tidymodels)
```

```{r setup, include=FALSE}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  
```

```{r}
ggplot(weather, aes(x = temp9am, y = temp3pm, color = location, shape = raintoday)) + 
  geom_point(size = 2.5, alpha = 0.8) +
  labs(
    title = "Morning vs. Afternoon Temperatures by Location",
    x = "Temperature at 9AM (°C)",
    y = "Temperature at 3PM (°C)",
    color = "Location",
    shape = "Rained Today",
    caption = "Source: ICA"
  ) +
  theme_minimal(base_size = 13)

```


