---
title: "Spatial Viz"
author: Phoebe Pan
date: 4/24/2025
format: 
  html:
    code-fold: true
    theme: cosmo
    toc: true
    number-sections: true
    warning: false
---

This section showcases my best spatial visualization, created and refined over the semester as I developed a stronger understanding of geographic data representation and interactive mapping techniques. The goal was to clearly communicate spatial relationships by integrating location-based data into a visual format that highlights geographic context.

The data comes from the our_fave_places.csv dataset, available at hash-mac.github.io, which includes favorite places submitted by students, along with their geographic coordinates. This dataset allowed me to build an interactive map using the leaflet package, enabling viewers to explore points of interest across the globe while reinforcing key skills in spatial data handling and visualization.

```{r, include=FALSE}
library(leaflet)
library(gplots)
library(rnaturalearth)
```

```{r, include=FALSE}
fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")
```

```{r}
leaflet(data = fave_places) |>
  addProviderTiles("USGS") |>
  addCircles(weight = 10, opacity = 1, color = col2hex("yellow")) |>
  addPolylines(
    lng = ~longitude,
    lat = ~latitude,
    color = col2hex("green")
  )
```

