---
title: "Univariate Viz"

author: Phoebe Pan
date: 4/24/2025
format: 
  html:
    code-fold: true
    theme: cosmo
    toc: true
    number-sections: true
---

This section showcases my best univariate visualizations, refined throughout the semester as I learned principles of effective design. The goal was to clearly communicate the distribution of a single variable using clean and professional plots.

The data comes Student Survey, available at hash-mac.github.io/stat112site-s25/data/survey.csv, which captures Macalester students’ outdoor preferences.

```{r setup, include=FALSE}
# Import data
# Used student survey data from ICA
survey <- read.csv("https://hash-mac.github.io/stat112site-s25/data/survey.csv")
library(ggplot2)
library(tidyverse)
```

```{r}
survey %>%
  count(hangout) %>%
  ggplot(aes(x = fct_reorder(hangout, n), y = n)) +
  geom_bar(stat = "identity", fill = "seagreen") +
  geom_text(aes(label = n), hjust = 2, vjust = -0.5) +
  labs(x = "Location", y = "Number of Students",
       title = "Where Students Prefer to Hang Out Outdoors") +
  coord_flip() +
  theme_minimal()
```
```{r}
ggplot(survey, aes(x = fav_temp_c)) +
  geom_histogram(binwidth = 2, fill = "coral", color = "white") +
  geom_vline(aes(xintercept = mean(fav_temp_c, na.rm = TRUE)), 
             linetype = "dashed", color = "steelblue", linewidth = 1) +
  annotate("text", 
           x = mean(survey$fav_temp_c, na.rm = TRUE) + 1.5, 
           y = 8, 
           label = "Mean", 
           color = "steelblue", 
           angle = 90, 
           vjust = -0.5, 
           size = 4) +
  labs(
    title = "Distribution of Preferred Outdoor Temperature",
    x = "Preferred Temperature (°C)",
    y = "Count"
  ) +
  coord_cartesian(xlim = c(-10, 40)) +
  theme_minimal(base_size = 14)
```


```{r}
ggplot(survey, aes(x = fav_temp_c)) +
  geom_density(fill = "coral", alpha = 0.5) +
  geom_vline(aes(xintercept = mean(fav_temp_c, na.rm = TRUE)),
             linetype = "dashed", color = "steelblue", linewidth = 1) +
  annotate("text",
           x = mean(survey$fav_temp_c, na.rm = TRUE) + 0.5,
           y = 0.055,
           label = "Mean",
           color = "steelblue",
           angle = 90,
           vjust = -0.5,
           size = 4) +
  geom_vline(aes(xintercept = median(fav_temp_c, na.rm = TRUE)),
             linetype = "dotted", color = "darkgreen", linewidth = 1) +
  annotate("text",
           x = median(survey$fav_temp_c, na.rm = TRUE) + 3,
           y = 0.055,
           label = "Median",
           color = "darkgreen",
           angle = 90,
           vjust = -0.5,
           size = 4) +
  labs(
    title = "Density of Student Temperature Preferences",
    x = "Preferred Temperature (°C)",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)

```


