---
title: "Exam_2"
author: Phoebe Pan
date: 4/24/2025
format: html
---
```{r, message=FALSE,warning=FALSE,echo=FALSE}
library(tidytuesdayR)
library(tidyverse)
library(ggplot2)
library(rnaturalearth)
library(dplyr)
library(sf)
```

### Data

**food_consumption.csv**

| variable      | class     | description                       |
|---------------|-----------|-----------------------------------|
| country       | character | Country Name                      |
| food_category | character | Food Category                     |
| consumption   | double    | Consumption (kg/person/year)      |
| co2_emmission | double    | CO2 Emission (kg CO2/person/year) |


```{r,echo = FALSE}
#| message: false
#| warning: false
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```

## Research Question: Which 5 countries consume the most food?

```{r}
top_consumers <- fc %>%
  group_by(country) %>%
  summarize(total_consumption = sum(consumption, na.rm = TRUE)) %>%
  arrange(desc(total_consumption)) %>%
  slice_head(n = 5)

top_consumers
```
The table clearly identifies the top five countries with the highest total food consumption (measured in kg/person/year). Finland ranks first with 639.8 kg/person/year, followed by Lithuania, Sweden, the Netherlands, and Albania. 

```{r fig.width=10, fig.height=10}
fc_country <- fc %>%
  group_by(country) %>%
  summarise(
    total_consumption = sum(consumption, na.rm = TRUE)
  ) %>%
  arrange(desc(total_consumption)) %>%
  mutate(color_group = ifelse(row_number() <= 5, "Top 5", "Others")) %>%
  slice(1:20) 

# Plot
ggplot(fc_country, aes(x = reorder(country, total_consumption), y = total_consumption, fill = color_group)) +
  geom_col() +
  scale_fill_manual(values = c("Top 5" = "red", "Others" = "steelblue")) +
  labs(
    title = "Top 20 Countries by Total Food Consumption",
    x = "Country",
    y = "Total Consumption (kg/person/year)",
    fill = NULL
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The bar plot visualizes total food consumption across the top 20 countries and highlights the top five in red. Finland prominently leading the chart.
