---
title: "Mapping the Museums of the United Kingdom"
format: html
---

## Intro

Museums are cultural landmarks, preserving the diverse history and heritage of the United Kingdom. This project explores the spatial distribution, types, and key characteristics of museums across England, Scotland, Wales, and Northern Ireland.
By leveraging open data from TidyTuesday, I visualize the UK’s museum landscape and examine how different governance models, subject matters, and locations shape the museum sector.
Through this analysis, I hope to contribute to a better understanding of the cultural infrastructure of the UK and highlight the important role museums play in connecting communities with their past.

---

## Data Acquisition and Preparation

For this analysis, I’m using data from the TidyTuesday project in 2022 (https://github.com/rfordatascience/tidytuesday/tree/main/data/2022/2022-11-22), which is sourced from Museums UK, providing details on thousands of museums, their locations, categories, and governance.

```{r}
# Load required packages
library(tidyverse)
library(sf)
library(ggplot2)
library(maps)
library(rnaturalearth)
library(RColorBrewer)
library(mosaic)
```

```{r}
# Import data
museums <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-11-22/museums.csv") 
```
Let's take a quick look at the data, it includes 35 variables -- we may want to clean the data to keep the tidiness.
```{r}
# Inspect data
head(museums)
```
```{r}
# data cleaning
museums_clean <- museums %>% 
  select(Name_of_museum,Latitude,Longitude,Admin_area, Accreditation,Governance,Size,Subject_Matter,Year_closed) %>% # keep only useful variables
  filter(Year_closed == "9999:9999") %>% # only look at the open museums
  mutate(country = str_trim(str_split(Admin_area, "/", simplify = TRUE)[,2])) %>% 
  mutate(subject = str_trim(str_split(Subject_Matter, "-", simplify = TRUE)[,1])) %>% 
  mutate(gov_cat = str_trim(str_split(Governance, "-", simplify = TRUE)[,1])) %>% 
  select(-Admin_area,-Subject_Matter,-Year_closed) # delete the variables that we've already organized
```

```{r}
# prpare the map background
uk_countries <- ne_states(country = "United Kingdom", returnclass = "sf")
```

---

## Analysis
```{r}
ggplot(museums_clean, aes(x = country, fill = Size)) +
  geom_bar()  +
  labs(x = "UK Country", y = "Count") +
  theme_minimal()

ggplot(museums_clean , aes(x = Size, fill = country)) +
  geom_bar(position = "fill") +
  labs(x = "Museum Size", y = "Count") +
  theme_minimal()
```
From the 2 plots above, we can observe that most UK museums are located in England, which has by far the largest number of museums in every size category. Scotland has the second highest count, followed by Wales and Northern Ireland, with the Channel Islands and Isle of Man having very few. Across all countries, small museums are the most common, while “huge” museums are rare everywhere.

```{r}
ggplot() +
  geom_sf(data = uk_countries, fill = "gray95", color = "gray70") +
  geom_point(
    data = museums_clean,
    aes(x = Longitude, y = Latitude, color = Size),
    size = 1.1, alpha = 0.2
  ) +
  coord_sf(xlim = c(-8, 2), ylim = c(49, 61)) +
  theme_void()
```
The map confirms our understanding about distribution of museums of different sizes -- the majority of them are small. Now let's shift to the the distribution of governance category by taking a quick look at the 2nd map. 


```{r}
ggplot() +
  geom_sf(data = uk_countries, fill = "gray95", color = "gray70") +
  geom_point(
    data = museums_clean,
    aes(x = Longitude, y = Latitude, color = gov_cat), 
    size = 1.1, alpha = 0.5
  ) +
  coord_sf(xlim = c(-8, 2), ylim = c(49, 61)) +
  theme_void()
```
We can see that the independent governance dominates the UK museum landscape, with government museums making up much smaller shares. Museums are most densely clustered in southern England but are spread across the country, including remote regions. In the following steps we would analyze the museum's governance composition combining with the subgroups with bar plot. 

```{r}
ggplot(museums_clean, aes(x = gov_cat, fill = Governance)) +
  geom_bar() +
  theme_minimal()

gov_counts <- museums_clean %>%
  count(Governance) %>%
  arrange(n) %>%
  mutate(Governance = factor(Governance, levels = Governance)) 


ggplot(gov_counts, aes(x = Governance, y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), vjust = -0.3, size = 3) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(
    x = "Governance Type",
    y = "Number of Museums",
    title = "Number of Museums by Governance Type"
  )
```
The previous geographic map shows that museums are densely clustered in southern England, and independent museums dominate the landscape, particularly in England. This pattern is clearly reflected in both bar charts, too. 

Meanwhile, both bar charts show that “Independent-Not_for_profit” and “Independent-Private” are by far the most common categories, together making up the majority of museums. Local authority-run museums are the main form of government governance, but they are much less numerous than independent ones.

Now, let's focus on the most common governance types and the size composition within each type.

```{r}
ggplot(museums_clean %>% filter(Governance %in% c("Independent-National_Trust", "Independent-Private","Government-Local_Authority","Independent-Not_for_profit")), aes(x = Governance, fill = Size)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(
    x = "Governance Type",
    y = "Number of Museums",
    title = "Number of Museums by Governance Type"
  )
```
Overall, small and medium-sized museums dominate each governance type, but government-managed museums are more likely to be medium or large compared to independents. This pattern suggests that independent museums are more numerous but generally smaller, while government-run museums, though fewer, tend to be larger institutions.

This distribution suggests the UK government prioritizes supporting fewer, larger museums with broad public impact. Meanwhile, the independent and nonprofit sector are encouraged to play a role in operating the many smaller, local museums, reflecting strong grassroots involvement and cultural diversity.



---

## Conclusion

My analysis highlights the rich and varied landscape of museums across the United Kingdom. The overwhelming presence of independent museums—especially small, not-for-profit, and private institutions—underscores the vital role of local communities and grassroots organizations in preserving and sharing culture. While England dominates in sheer number and variety, the contributions from Scotland, Wales, and Northern Ireland further illustrate the variety of regional heritage.

Government involvement, while less extensive in number, is clearly focused on supporting larger institutions with wider public reach. This targeted approach allows the government to maintain and promote key national assets, while also enabling independent organizations to flourish at the local level -- it has achieved its balance. 

Overall, museums remain crucial to connecting people with history, art, and science. The patterns revealed here point to a healthy ecosystem where both state and independent actors contribute tothe cultural heritage for future generations.

---

## Reference

TidyTuesday. (2022). Museums Dataset. Retrieved from 
https://github.com/rfordatascience/tidytuesday/tree/main/data/2022/2022-11-22













